---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

```{r}
library('maps') 
library('ggplot2') 
library('dplyr')
library(datasets)
```

# Assignment 1

**Collaborators: w/ Sophia Restaino **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.

### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
library(datasets)
```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

```{r}
dat <- USArrests
dat.untouched <- dat
```

It is useful to rename the dataset so that you are able to keep track of your own datasets that you will be working with on your project. When working with many datasets, it might get confusing regarding which version of the data you are using (ie. if there is any manipulation to the data involved). Best practices involve making an untouched version so you can relook at the original data. 

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=TRUE}
dat$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.

List the variables contained in the dataset `USArrests`.

```{r}
names(dat)
```

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`?
```{r}
?USArrests
```

Answer: Murder is a quantitative variable because it is a numerical value with units. It measures the number of murder arrests in each state (per 100,000) in 1973.

What R Type of variable is it?
```{r}
class(dat$Murder)
```

Answer: The R type of variable is "numeric" as shown by the output of the code above. 

### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

Answer: This dataset is looking at the types of arrests in 1973 per 100,000 people in the 50 states. There are 50 rows (the states) and 4 main variables we are looking at, including murder, assault, urban population percentages, and rape arrests. 

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
hist(dat$Murder,
     main = "Histogram of Murders",
     xlab = "Number of Murders",
     ylab = "Frequency",
     xlim = c(0,20),
     ylim = c(0,15))
```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
mean(dat$Murder)
median(dat$Murder)
summary(dat$Murder)
```

The mean shows the average amount of murders per 100,000 residents for each state. The median reports the number for the state that is in the middle of the other numbers in the dataset. Between the 1st quartile, median, 3rd quartile, and the max, we are able to see enough of the the data distribution that we need. 1st quartile shows that 25% of the state's numbers reported are below 4.075 and the 3rd quartile shows that the 50.1%-75% are between 7.25-11.25. 

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
max(dat$Assault)
hist(dat$Assault,
     main = "Histogram of Assault",
     xlab = "Number of Assaults",
     ylab = "Frequency",
     xlim = c(0,350),
     ylim = c(0,12))
```

```{r}
max(dat$Rape)
hist(dat$Rape,
     main = "Histogram of Rape",
     xlab = "Number of Rapes",
     ylab = "Frequency",
     xlim = c(0,50),
     ylim = c(0,12))
```

What does the command par do, in your own words (you can look this up by asking R `?par`)?
```{r}
?par
```

Answer: Par allows you to set graphical parameters. An example of using par can be seen below in which I used par to put graphs together to be seen side by side. You can use it to put, for example, 3 side by side or 3 on top of each other. 

What can you learn from plotting the histograms together?
```{r}
par(mfrow=c(3,1))

hist(dat$Murder,
     main = "Histogram of Murders",
     xlab = "Number of Murders",
     ylab = "Frequency",
     xlim = c(0,20),
     ylim = c(0,15))

hist(dat$Assault,
     main = "Histogram of Assault",
     xlab = "Number of Assaults",
     ylab = "Frequency",
     xlim = c(0,350),
     ylim = c(0,12))

hist(dat$Rape,
     main = "Histogram of Rape",
     xlab = "Number of Rapes",
     ylab = "Frequency",
     xlim = c(0,50),
     ylim = c(0,12))
```

Answer: You are able to compare the frequencies of the assaults, murders, and rape all togehter at once. It is an easy way to display the data. 
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval = FALSE, fig.width = 7.5, fig.height = 4}
ggplot(dat,
       aes(map_id=state, fill=Murder)) + 
       geom_map(map=map_data("state")) + 
       expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

Answer: We are making a map with the murder data, plotting the different murder rates across each state. The first line of this code is calling the maps package. The second line calls the ggplot2 package which allows you to make different kinds of graphs, similar to baseR plot functions. So, when making the map, "ggplot" is the function, "dat" is the dataset we are pulling information from, and "aes" is responsible for setting the aesthetics of the map we are making. "fill" is telling R that we want the aesthetics to be filled with the data from the "Murder" column. After this, the "+" tacks on the map that is actually being made, using the "state" data. After this, "expand_limits" helps us put the latitude and longitude ticks on the visualization, behind the map. 

# Assignment 2

### Problem 1: Load Data
Read in the data
```{r}
library(dplyr)
nsduh <- read.csv("dat.nsduh.small.1.csv")
nsduh.untouched <- nsduh
names(nsduh)
```

Dimensions of the dataset
```{r}
dim(nsduh)
```

## Problem 2: Variables
Describe the variables in the dataset.
What is this dataset about? Who collected the data, what kind of sample is it, and what was the purpose of generating the data?

This dataset is a survey collecting information on people's tendency/history of using marijuana, cigarettes, and alcohol. There are also some demographics/personal information collected, such as the respondent's age, sex, and sexual attraction. This information is all compiled by the National Survey on Drug Use and Health in order to analyze substance use across the nation. 

## Problem 3: Age and gender
What is the age distribution of the sample like? Make sure you read the codebook to know what the variable values mean.
Do you think this age distribution representative of the US population? Why or why not?
```{r}
class(nsduh$age2)
mean(nsduh$age2)
median(nsduh$age2)
hist(nsduh$age2,
     main = "Histogram of Ages",
     xlab = "Age Buckets",
     ylab = "Frequency",
     xlim = c(0,20))
```

According to this histogram, along with the information on the value meanings in the codebook, most of the respondents for this survey answered 15. This indicates that they are between 35-49 years old. This survey took people who were at least 12 years old, but compared to the other age buckets, younger people were not as represented. This makes sense, as the data was collected through a survey, so slightly older people were probably more likely to interact with it. I think this is somewhat reflective of the national population, but I would probably want a larger sample size just to be as accurate as possible. A quick google search shows that in 2019, the average age of a person was 38.1 years old. The average age of respondents in this survey was roughly between 26-29 years old. Thus, this information shows that the survey is slightly on the younger side and could use more respondents to be more reflective of the national population. 

Is the sample balanced in terms of gender? If not, are there more females or males?
```{r}
nsduh %>%
  count(irsex)
```
It is decently balanced. There are 91 males and 80 females. 

Use this code to draw a stacked bar plot to view the relationship between sex and age. What can you conclude from this plot?
```{r}
tab.agesex <- table(nsduh$irsex, nsduh$age2)
min(nsduh$age2)
barplot(tab.agesex,
        main = "Stacked barchart",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.agesex),
        beside = FALSE) # Stacked bars (default)
```
This barchart allows us to see the gender breakdown per age bucket in the survey. For example, we are able to see that for those who are in the "15" age bucket, the respondents are pretty evenly split between males and females. We are also able to see instances in which the age buckets are heavily dominated by female or male respondents. For the 17 and 18 year olds (buckets 6 nd 7), according to the graph, most respondents also reported that they are male. For the 19 year old respondents (bucket 8), many reported that they are female, rather than males. Overall, most of the age buckets are decently even between female and male respondents. 

## Problem 4: Substance use
For which of the three substances included in the dataset (marijuana, alcohol, and cigarettes) do individuals tend to use the substance earlier?
```{r}
sum(min(nsduh$mjage))
sum(min(nsduh$cigage))
sum(min(nsduh$iralcage))
```
Individuals tend to use alcohol at an earlier age than compared to when they would use marijuana or cigarettes. 

## Problem 5: Sexual attraction
What does the distribution of sexual attraction look like? Is this what you expected?
```{r}
nsduh$sexna <- nsduh$sexatract
nsduh$sexna[nsduh$sexatract == 85 | nsduh$sexatract == 94 | nsduh$sexatract == 97 |
            nsduh$sexatract == 98 | nsduh$sexatract == 99] <- NA 
length(nsduh$sexna[nsduh$sexna == "NA"])
hist(nsduh$sexna,
     main = "Histogram of Sexual Attraction",
     xlab = "Age Buckets",
     ylab = "Frequency",
     xlim = c(1,6))
```
First, I excluded people on the survey who skipped this question, did not answer, didn't know, or refused this question. I found that in this dataset, this only applied to 3 people, so I felt comfortable excluding them without much harm to the strength of the data. An overwhelming majority of the people who took this survey are straight, as in they are only attracted to people of the opposite sex. The second most popular response was that people are mostly attracted to people of the opposite sex. People who are bi was the next most common answer, followed by people who are mostly attracted to the same sex, and finally people who are only attracted to the same sex. The least common response was people who are not sure. This is what I expected, and I think this data is representative of the greater population. 

What is the distribution of sexual attraction by gender? 
```{r}
tab.gensex <- table(nsduh$irsex, nsduh$sexna)
barplot(tab.gensex,
        main = "Stacked barchart",
        xlab = "Sexual Attraction Category", ylab = "Frequency",
        legend.text = rownames(tab.gensex),
        beside = FALSE) # Stacked bars (default)
```
Most females and males who took the survey are straight. Women are more likely to report that they are mostly attracted to the opposite sex and are equally attracted to males and females than the other male respondents. 

## Problem 6: English speaking
What does the distribution of English speaking look like in the sample? Is this what you might expect for a random sample of the US population?
```{r}
speakengl <- nsduh$speakengl
engl.freq <- table(speakengl)
barplot(engl.freq,
        ylim = c(0,200))
```
Most of the people who took the survey are able to speak english very well. This is followed by those who said they can speak english well, followed by those who said they can't speak english very well. This is not surprising as a random sample of the US because most people are able to speak english very well; however, there are people who can speak english, but not as their first language. This might cause their english to not be as strong, so it makes sense that the survey responses are reflective of this factor. 

Are there more English speaker females or males?
```{r}
tab.genengl <- table(nsduh$speakengl, nsduh$irsex)
barplot(tab.genengl,
        main = "Stacked barchart",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.genengl),
        beside = FALSE,
        ylim = c(0,100)) # Stacked bars (default)
```
There are more english speaking males. This might be caused by a few factors with one of the more obvious being that more males took the survey than females, so it makes sense that, in looking at pure frequency, they might have more people answer that they can speak english well. 

# Assignment 3
```{r}
setwd("~/Desktop/Fall 2021/CRIM 250")
library(readr)
library(knitr)
library(ggplot2)
dat.crime <- read_delim("crime_simple.txt", delim = "\t")
```

This is a dataset from a textbook by Brian S. Everitt about crime in the US in 1960. The data originate from the Uniform Crime Report of the FBI and other government sources. The data for 47 states of the USA are given. 

Here is the codebook:

R: Crime rate: # of offenses reported to police per million population

Age: The number of males of age 14-24 per 1000 population

S: Indicator variable for Southern states (0 = No, 1 = Yes)

Ed: Mean of years of schooling x 10 for persons of age 25 or older

Ex0: 1960 per capita expenditure on police by state and local government

Ex1: 1959 per capita expenditure on police by state and local government

LF: Labor force participation rate per 1000 civilian urban males age 14-24

M: The number of males per 1000 females

N: State population size in hundred thousands

NW: The number of non-whites per 1000 population

U1: Unemployment rate of urban males per 1000 of age 14-24

U2: Unemployment rate of urban males per 1000 of age 35-39

W: Median value of transferable goods and assets or family income in tens of $

X: The number of families per 1000 earning below 1/2 the median income


We are interested in checking whether the reported crime rate (# of offenses reported to police per million population) and the average education (mean number of years of schooling for persons of age 25 or older) are related. 


1. How many observations are there in the dataset? To what does each observation correspond?

```{r}
dim(dat.crime)
```
There are 47 observations in this dataset, and each of the rows represent one of 47 states included in this dataset. 

2. Draw a scatterplot of the two variables. Calculate the correlation between the two variables. Can you come up with an explanation for this relationship?

```{r, fig.width=6, fig.height=4}
ggplot(dat.crime, aes(x = Ed, y = U2)) + 
    geom_point() +
    geom_smooth(method = lm, se=FALSE)

cor(dat.crime$Ed,
    dat.crime$U2,
    method = "pearson",
    use = "complete.obs")
```
There is a slight negative correlation between the variables, and when calculating the correlation using the pearson method, it shows that it comes out to -0.22. This makes sense because as education levels increase, then crime rates will likely decrease. As education decreases, then crime rates tend to increase. 

3. Regress reported crime rate (y) on average education (x) and call this linear model `crime.lm` and write the summary of the regression by using this code, which makes it look a little nicer `{r, eval=FALSE} kable(summary(crime.lm)$coef, digits = 2)`.

```{r} 
crime.lm <- lm(R ~ Ed, data = dat.crime)
summary(crime.lm)
```

4. Are the four assumptions of linear regression satisfied? To answer this, draw the relevant plots. (Write a maximum of one sentence per assumption.)

```{r} 
# Linearity
x <- dat.crime$Ed
y <- dat.crime$R

## residuals vs x
plot(x, y, main = "Relationship between Crime Rate and Education",
     xlab = "Education", ylab = "Crime Rate",
     pch = 19, frame = FALSE)
abline(crime.lm, col = "blue")

## residuals vs fitted
plot(crime.lm, which=1)

# Normal Population
## qq plot
plot(crime.lm, which=2)

# Homoscedasticity
## scale location plot
plot(crime.lm, which=3)
```

Linearity assumption: This is satisfied because the scatterplot abline shows a consistent linear relationship between the variables, and the red line in the residuals vs fitted graph is flat which indicates there is no discernible non-linear trend to the residuals.
Equal variation assumption (homoscedasticity): Homoscedasticity looks at each point's distance away from the line, and most of the dots are constantly the same distance for their respective points.
Normal population assumption: The QQ plot shows that most of the points follow the behavior of the line which indicates that this is mostly normal. 
Independence assumption: Independence assumption is satisfied because there are no patterns in the scatterplot showing the relationship between R and Ed variables. 

5. Is the relationship between reported crime and average education statistically significant? Report the estimated coefficient of the slope, the standard error, and the p-value. What does it mean for the relationship to be statistically significant?

```{r}
summary(crime.lm)
```
Looking at the output from the summary of the regression, we can see that the p-value is less than 0.05 which indicates statistical significance. This means that given the relationship of crime reported and years of education, we fail to reject the null hypothesis with 95% confidence. 

6. How are reported crime and average education related? In other words, for every unit increase in average education, how does reported crime rate change (per million) per state?

This data shows that with every unit increase in education, there is a (positive) 1.116 unit increase in the likeliness that reported crime increases as well. This is surprising given that I would think that with increase in education, there is a decrease in crime, but this is not what the data shows. 

7. Can you conclude that if individuals were to receive more education, then crime will be reported more often? Why or why not?

One cannot draw this conclusion. First, the codebook doesn't draw a differentiation between crime rate and numbers reported. Our regression does not take into account enough factors to draw a strong conclusion on this relationship, and there aren't time series data available for us to look at changes.

# Assignment 4

```{r}
# 3 Data Visualization

# install.packages("tidyverse")
library(tidyverse)
## installing and loading the package "tidyverse" from the library

mpg
## printing a bit of the dataset into the console

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
## creating a basic ggplot with the datapoints from mpg

# ggplot(data = <DATA>) + 
#  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
## creating a reusable template for future graphs

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
## mapping the same basic graph with mpg data but changing the color to be grouped by class
## this happens in the aes with "color"

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
## mapping the same basic graph but with changes to the size of the point
## "size" of the dot is grouped by class

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
## mapping the same basic graph but with changes to the transparency level of the points
## in the aes with "alpha"

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
## mapping the same basic graph but with changes to the shape of the points
## in the aes with "shape"

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
## mapping the same basic graph but changing the color of the dots manually
## see: not in the aes

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
## points will not be blue because "color" is in the aes

# ggplot(data = mpg) 
# + geom_point(mapping = aes(x = displ, y = hwy))
## the "+" needs to be at the end of the line, not at the beginning
## important syntax!

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
## same basic graph but split into "facets" which displays different sets of subplots
## wrapped by class, and with 2 rows

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
## facets the graphs by 2 variables

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
## manually setting the x/y to be drv and cyl
## not a facet graph so not subplots of these together

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
## same basic graph split by variable "drv" but the graphs are displayed horizontally bc of the "."

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
## same basic graph split by variable "cyl" but the graphs are displayed vertically bc of the "."

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
## same as line 56 code

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
## geom point graph

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
## same graph as immediately above, but changes the shape of how the data is displayed by using "geom_smooth"

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
## basic "geom_smooth" graph but with linetype based on "drv" variable

ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE)
## adds on to graph immediately above, changing the color to match the linetype grouping

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
## plots both ways, geom_bar and geom_smooth on one graph

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
## same as graph above

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
## same as graph above but with color changes for the points on the graph, based on class

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
## essentially the same as the graph above, but with changes to the geom_smooth line
## changes include only graphing the subcompact class

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
## graphs grouped by drv with colors for line and points corresponding

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
## the above 2 do exactly the same, just a matter of where you are putting the command information

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
## basic ggplot bar

ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
## same as graph above, "stat_count" is interchangeable with "geom_bar"

demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
## creates a dataset to manually change the height of the blocks on the geom_bar

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
## creates basic graph using above data "demo"

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
## geom_bar using proportions, rather than using the raw data

ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median)
## summarizes the y values for each x value and creates a graph accordingly

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
## colors the bars in the aes

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
## combines cut and clarity in order to fill the blocks with different colors

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
## different ways of playing around with transparency, fill, outlines in color, etc.

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
## basic geom_bar but with different fill and position
## fill makes all of the blocks the same height
## easier to compare proportions

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
## puts overlapping objects right next to each other, rather than on top of each other

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
## adds "noise" to each point, which can either improve your graph or not

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
## creates basic boxplot
## same thing but the coord_flip changes the x/y

nz <- map_data("nz")
## loads nz map data

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")
## basic mapping of nz

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
## corrects the ratio on the map above

bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
## creates data

bar + coord_flip()
## flips cartesian coordinates to plot

bar + coord_polar()
## uses polar coordinates to create visualization

# ggplot(data = <DATA>) + 
  # <GEOM_FUNCTION>(
  #  mapping = aes(<MAPPINGS>),
  #  stat = <STAT>, 
  #  position = <POSITION>
  # ) +
  # <COORDINATE_FUNCTION> +
  # <FACET_FUNCTION>
## template

# 28 Graphics for Communication

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
## adds title to geom plot we created before

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov")
## adds caption and subtitle to graphs that we created above

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type")
## adds x and y label to the axes and names the label

df <- tibble(
  x = runif(10),
  y = runif(10))
## creates data frame using tibble

ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + beta + frac(delta, theta)))
## plots the data above using geom_point and adds the mathematic equations corresponding with the axes

best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)
## creates dataset based on mpg with specific qualifications for grouping and filters
## uses pipe %

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
## graphs mpg data and layers on text dependent on the best_in_class data we just made
## adds text on the graph corresponding with the points

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
## alters the way that the text is presented on the graph (transparency, and adds box around text)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
## ensures that the labels we just added are not overlapping on the graph

class_avg <- mpg %>%
  group_by(class) %>%
  summarise(
    displ = median(displ),
    hwy = median(hwy))
ggplot(mpg, aes(displ, hwy, colour = class)) +
  ggrepel::geom_label_repel(aes(label = class),
                            data = class_avg,
                            size = 6,
                            label.size = 0,
                            segment.color = NA) +
  geom_point() +
  theme(legend.position = "none")
## creating a dataset of mpg with certain characteristics
## making a ggplot with the colors, labels, etc. corresponding to the class
## specifying no labels

label <- mpg %>%
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \nrelated to decreasing fuel economy.")
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
## creating "label" to add to the graph aes and placing it in the top right

label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy.")
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
## "tibble" creates the data fram and then adding that to the graph
## essentially the same as the above functions

"Increasing engine size is related to decreasing fuel economy." %>%
  stringr::str_wrap(width = 40) %>%
  writeLines()
## another way of breaking up the text, rather than using"\n"

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
## basic plot

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_discrete()
## showing the default scales that ggplot adds

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
## adding our own scales

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
## suppressing all the labels together (doesn't share absolute numbers)

presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y")
## graph with things to help us control the labels for this graph using "breaks"

base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right")
## base + the theme that will indicate where the legend is on the graph
## how to manually change that from the default (which the the right)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))
## adding a theme and the color guide
## overrides some of the defaults for sizing and color

ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()
## different graph with geom_bin2d

ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
## log version of graph above

ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
## adding x/y scales to log graph above

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
## adding color palette

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set1")
## color palette based on drv variable

presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
## adding colors to match political party

df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000))
ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()
ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
## more colors

ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))
mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
## zooming in on the points from the graph above

suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")
## creating dataframes

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point()
ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point()
## plotting the dataframes basing color on drv variable

x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_colour_discrete(limits = unique(mpg$drv))
## limiting the full data to make scales

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
## using the scales we just made to plot the respective suv and compact dataframes

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_bw()
## adding a theme to the plot

ggplot(mpg, aes(displ, hwy)) + geom_point()
ggsave("my-plot.pdf")
## saving the plot above with ggsave
```

